Inference of proto-neutron star properties from gravitational-wave data in core-collapse supernovae

### 2021-02-02 conclusion.tex
version corresponding to PRD resubmission

In the present study we have assumed that the sources are optimally oriented. The reported distances at which
we can infer the time evolution of $M_{\rm PNS}/R_{\rm PNS}^2$ must thus be regarded as upper limits. Those figures may decrease by a factor 2--3 on average for a source located anywhere in the sky. Furthermore, we have not considered the detectability prospects of  CCSN waveforms in the realistic case in which the  interferometers operated as a detector network. 
We defer an improved implementation of our approach for a forthcoming publication. Finally, we note that the method discussed in this work can be adapted to other PNS oscillation modes, by simply changing a few parameters such as the initial frequency range of the mode and its monotonic raise or descent. Being able to reconstruct several modes in the same GW signal would potentially allow to individually infer the mass and the radius of the PNS in core-collapse supernova explosions.

>>>>

In the present study we have assumed that the sources are optimally oriented. The reported distances at which
we can infer the time evolution of $M_{\rm PNS}/R_{\rm PNS}^2$ must thus be regarded as upper limits. Those figures may decrease by a factor 2--3 on average for a source located anywhere in the sky.
Moreover, we have used 2D CCSN simulations that are known to produce higher amplitude GW signals than 3D CCSN simulations~\citep{Radice:2018usf, Mezzacappa:2020lsn}. This fact and the low frequency standing accretion shock instability that appears only in 3D simulations may reduce further the performance of the method.
Furthermore, we have not considered the detectability prospects of  CCSN waveforms in the realistic case in which the  interferometers operated as a detector network. 
We defer an improved implementation of our approach for a forthcoming publication. Finally, we note that the method discussed in this work can be adapted to other PNS oscillation modes, by simply changing a few parameters such as the initial frequency range of the mode and its monotonic raise or descent. Being able to reconstruct several modes in the same GW signal would potentially allow to individually infer the mass and the radius of the PNS in core-collapse supernova explosions.

### 2021-02-02 introduction.tex
version corresponding to PRD resubmission

{In this work we introduce a method to infer PNS properties, namely a combination of the mass and radius, using GW information. For this purpose we have developed an algorithm to  
  extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN. This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in~\citep{Torres:2019b} (different authors may have slightly different naming convention). Next, we use the universal relations obtained by~\citep{Torres:2019b}, based on a set of 1D simulations, to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ (the PNS surface gravity), \textcolor{red}{where $M_{\rm PNS}$ is the mass of the PNS and $R_{\rm PNS}$ is the radius where the density is lower than \unit[$10^{12}$]{$\rm g/cm^3$}}.
%  where  $M_{\rm PNS}$ and $R_{\rm PNS}$ are the mass and the radius of the PNS.
Finally, using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.

>>>>

{In this work we introduce a method to infer PNS properties, namely a combination of the mass and radius, using GW information. For this purpose we have developed an algorithm to  
  extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN. This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in~\citep{Torres:2019b} 
 %(different authors may have slightly different naming convention)
Next, we use the universal relations obtained by~\citep{Torres:2019b}, based on a set of 1D simulations, to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ (the PNS surface gravity), where $M_{\rm PNS}$ is the mass of the PNS and $R_{\rm PNS}$ is the radius where the density is lower than \unit[$10^{11}$]{$\rm g/cm^3$}.
%  where  $M_{\rm PNS}$ and $R_{\rm PNS}$ are the mass and the radius of the PNS.
This same mode has been interpreted in some works as a f or a p-mode
\citep[e.g.][]{Sotani:2016,Radice:2018usf,Mezzacappa:2020lsn}. 
However, this does not affects our analysis, which 
only uses the relation between the frequency of the main mode emitting in GWs (regardless
of its nature) and the properties of the PNS.  
Finally, using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.

### 2021-02-02 methods.tex
version corresponding to PRD resubmission

\begin{figure}[t]
 \centering
 \includegraphics[width=0.45\textwidth]{plots/model}
 \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from our 18 1D simulations of the model set. The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the 18 simulation data points. \textcolor{red}{We have not made the distinction between the different simulations since we are only interested in the relationship between the variables.} }
 \label{fig:LMVAR}
\end{figure}

@@

\textcolor{red}{For the mean of the ratios $r_i$, the cubic model has the best fitting amongst polynomials of lower degrees, and for the mean of the log standard deviation $\log\sigma_i$, the quadratic model has the best fit compared to a polynomial of degree one.  The models were compared according to the Akaike information criterion.  The estimated coefficients are given in Table \ref{tab:model} and the data and fit of the final model including 95\% confidence bands are displayed in
Figure~\ref{fig:LMVAR}.}

>>>>

\begin{figure}[t]
 \centering
 \includegraphics[width=0.45\textwidth]{plots/model}
 \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from our 18 1D simulations of the model set. The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the 18 simulation data points. We have not made the distinction between the different simulations since we are only interested in the relationship between the variables.}
 \label{fig:LMVAR}
\end{figure}

@@

For the mean of the ratios $r_i$, the cubic model has the best fitting amongst polynomials of lower degrees, and for the mean of the log standard deviation $\log\sigma_i$, the quadratic model has the best fit compared to a polynomial of degree one. The models were compared according to the Akaike information criterion.  The estimated coefficients are given in Table \ref{tab:model} and the data and fit of the final model, including 95\% confidence bands, are displayed in Figure~\ref{fig:LMVAR}.
